DOpElib - Deal Optimization Environment

### 2012-10-20 Examples/Documentation/foreword.tex
deal.II -> Differential Equations and

\chapter{Foreword}
In this report, we describe the DOpE
\textit{Deal Optimization Environment} project. 
Originally, the project was initiated in the year 2009 at the 
University of Heidelberg (Germany) in the numerical analysis
group of Rolf Rannacher.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{Deal Optimization Environment} (DOpE) project is 
based on the \textit{deal.II} finite element library which has been developed
 initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
Its main feature is to give a unified interface to high level algorithms such as 
time stepping methods, nonlinear solvers and optimization routines. 
We aim that the user should only need to write those parts
of the code that are problem dependent while all invariant parts of the algorithms
should be reusable without any need for further coding.
In particular, the user should be able to switch between various different 
algorithms without the need to rewrite the problem dependent code, though he or she will
have to replace the algorithm object with an other one.

>>>>

\chapter{Foreword}
In this report, we describe the DOpE
\textit{Differential Equations and  Optimization Environment} project. 
Originally, the project was initiated in the year 2009 at the 
University of Heidelberg (Germany) in the numerical analysis
group of Rolf Rannacher.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The \textit{Differential Equations and  Optimization Environment} (DOpE) project is 
mainly based on the \textit{deal.II} finite element library which has been developed
 initially by W. Bangerth, R. Hartmann, and G. Kanschat \cite{deal}.
Its main feature is to give a unified interface to high level algorithms such as 
time stepping methods, nonlinear solvers and optimization routines. 
We aim that the user should only need to write those parts
of the code that are problem dependent while all invariant parts of the algorithms
should be reusable without any need for further coding.
In particular, the user should be able to switch between various different 
algorithms without the need to rewrite the problem dependent code, though he or she will
have to replace the algorithm object with an other one.

### 2012-10-20 Examples/Documentation/introduction.tex
Update of the documentation, includes now higher order mappings.
deal.II -> Differential Equations and

\subsection{deal.II}
This library is based upon \texttt{deal.II} hence in order to run 
\texttt{DOpElib} you need a running copy of \texttt{deal.II}.

@@

\subsubsection{Auto-generated Problem Descriptions}
\begin{itemize}
  \item \texttt{problemdata/stateproblem.h} This is the problem description for the (forward/primal) PDE constraint.
    {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
  \item \texttt{problemdata/initialproblem.h} This is the problem descriptor to compute the finite element representation
    of the initial values. This is generated by the different time-stepping schemes based upon the defined 
    representation by the pde, which is set to the component wise $L^2$ projection by default.
  \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 
    forward/primal PDE.
  \item \texttt{problemdata/tangentdirichletdata.h} This class contains the Dirichlet data for the tangent PDE, i.e.,
    the first derivative of the Dirichlet data.
  \item \texttt{problemdata/transposedgradientdirichletdata.h} This contains the transposed Dirichlet data needed 
    to calculate the gradient of the reduced objective functional, 
    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
  \item \texttt{problemdata/transposedhessiandirichletdata.h} This contains the transposed Dirichlet data needed 
    to calculate the hessian of the reduced objective functional, 
    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
\end{itemize}
\subsubsection{Management of Time Dependent Problems}
\begin{itemize}
\item \texttt{include/timedofhandler.h} DoFHandler responsible for the management of the timedofs (this is a part of the \texttt{SpaceTimeDoFHandler}-classes). Basically a wrapper for a $1d$ \texttt{deal.II}-DoFHandler.
\item \texttt{include/timeiterator.h} This class works as an iterator on the \texttt{TimeDoFHandler}.                 
\end{itemize}
\subsection{Wrapper classes}
\begin{itemize}
  \item \texttt{wrapper/dofhandler\underline{ }wrapper.h} A wrapper class for the \texttt{deal.II} DoFHandlers. This 
    class is needed to provide support for the \texttt{dim $= 0$} case and to have a uniform interface to 
    DoFHandler and HPDoFHandler.
  \item \texttt{wrapper/preconditioner\underline{ }wrapper.h} Contains wrappers for several of the preconditioners
    in \texttt{deal.II}. This is required since unfortunately the preconditioners in \texttt{deal.II} have different
    interfaces for their initialization.
  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
    not only on space but also on time.
  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
    wrapper to the one provided by SNOPT to allow automatic construction of the functions required 
    by SNOPT using our library.
  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
\end{itemize}

>>>>

\subsection{deal.II}
This library is mainly based upon \texttt{deal.II} hence in order to run 
\texttt{DOpElib} you need a running copy of \texttt{deal.II}.

@@

\subsubsection{Auto-generated Problem Descriptions}
\begin{itemize}
  \item \texttt{problemdata/stateproblem.h} This is the problem description for the (forward/primal) PDE constraint.
    {\bf Similar descriptors will be build for the other problems (adjoint, tangent, $\ldots$) when time allows.}
  \item \texttt{problemdata/initialproblem.h} This is the problem descriptor to compute the finite element representation
    of the initial values. This is generated by the different time-stepping schemes based upon the defined 
    representation by the pde, which is set to the component wise $L^2$ projection by default.
  \item \texttt{problemdata/primaldirichletdata.h} This class contains the Dirichlet data for the 
    forward/primal PDE.
  \item \texttt{problemdata/tangentdirichletdata.h} This class contains the Dirichlet data for the tangent PDE, i.e.,
    the first derivative of the Dirichlet data.
  \item \texttt{problemdata/transposedgradientdirichletdata.h} This contains the transposed Dirichlet data needed 
    to calculate the gradient of the reduced objective functional, 
    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
  \item \texttt{problemdata/transposedhessiandirichletdata.h} This contains the transposed Dirichlet data needed 
    to calculate the hessian of the reduced objective functional, 
    for detail see \texttt{interfaces/transposeddirichletdatainterface.h}.
\end{itemize}
\subsubsection{Management of Time Dependent Problems}
\begin{itemize}
\item \texttt{include/timedofhandler.h} DoFHandler responsible for the management of the timedofs (this is a part of the \texttt{SpaceTimeDoFHandler}-classes). Basically a wrapper for a $1d$ \texttt{deal.II}-DoFHandler.
\item \texttt{include/timeiterator.h} This class works as an iterator on the \texttt{TimeDoFHandler}.                 
\end{itemize}
\subsection{Wrapper classes}
\begin{itemize}
  \item \texttt{wrapper/dofhandler\underline{ }wrapper.h} A wrapper class for the \texttt{deal.II} DoFHandlers. This 
    class is needed to provide support for the \texttt{dim $= 0$} case and to have a uniform interface to 
    DoFHandler and HPDoFHandler.
  \item \texttt{wrapper/fevalues\underline{ }wrapper.h} {\bf Will be removed soon!}
  \item \texttt{wrapper/function\underline{ }wrapper.h} An interface that allows to use functions that depend 
    not only on space but also on time.
  \item \texttt{wrapper/mapping\underline{ }wrapper.h} An interface that allows to use \texttt{deal.II}-mappings as well as  \texttt{deal.II}-mapping collections depending of the DoFHandler in use. To this end, the class has a template parameter \texttt{DOFHANDLER}.
  \item \texttt{wrapper/preconditioner\underline{ }wrapper.h} Contains wrappers for several of the preconditioners
    in \texttt{deal.II}. This is required since unfortunately the preconditioners in \texttt{deal.II} have different
    interfaces for their initialization.
  \item \texttt{wrapper/snopt\underline{ }wrapper.h} An interface to the SNOPT fortran library. This is an additional
    wrapper to the one provided by SNOPT to allow automatic construction of the functions required 
    by SNOPT using our library.
\end{itemize}
